---
title: "Fine scale SCTLD Lower FRT: Statistical Analyses"
output: html_notebook
---

## This notebook includes:
- Coral cover analysis: aligned ranks transformation 2-way ANOVA (cover_long.csv)
- Percent coral tissue loss per week per colony analysis (SWG_SCTLDprogrates.csv):
    - kruskal-wallis and barplot for differences among sites
    - species differences, kruskal-wallis test
- Temperature correlations: Pearson cor tests (extended_envdisFig3.csv)
- Bleaching and SCTLD: Chi-square tests (SCTLD_END_Vpub_ts.csv)
- Are certain species more likely to show signs of SCTLD: fishers (SCTLD_END_Vpub.csv)
- Size and SCTLD: wilcox test for just 11 susceptible species (SCTLD_END_Vpub.csv) and then multiple comparisons wiht correction for within species
- Time of disease onset: bar graphs (SCTLD_END_Vpub.csv,timedif_surveys.csv)
- Reef indices of susceptibility
    - linear regressions for prevalence & density, initial coral cover, shannon diversity,and size
    - PCoA for species at each quadrat in relation to prevalence

### Set up and load in packages you need
```{r, message=FALSE, warning=FALSE}
library(abind)
library(sciplot)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(ggpubr)
library(vegan)
library(nlme)
library(car)
library(patchwork)
library(ARTool)
library(phia)
getwd()
```

Version control: R version 4.0.2 (2020-06-22)
```{r}
sessionInfo()$R.version
```


#### Commonly used custom functions for running summary statistics

```{r}

my.se<-function(x){
  sd(x,na.rm=TRUE)/sqrt(length(x))
}
my.se.rows<-function(x){
  se.tp<-c()
  for(i in 1:nrow(x)){
    se.tp[i]<-my.se(x[i,1:2])
  }
  return(se.tp)
}
my.se.cols<-function(x,a){
  se.tp<-c()
  for (i in a:ncol(x)){
      se.tp[i]<-my.se(x[,i])
  }
  return(se.tp[12:32])
}

```

#### Load in data
```{r}
my.data<-read.csv("SCTLD_END_Vpub.csv") 
colnames(my.data)[1]<-"Site"
nrow(my.data)

# make a column with factors for if the coral got disease or not throughout the entire survey
my.data$glom<-as.factor(my.data$total_bin) 
my.data$glom<-revalue(my.data$glom, c("0"="Healthy", "1"="SCTLD")) 
head(my.data$glom)

# subset into sites
## site1 is midchannel, site2 is offshore, and site3 is nearshore
site1<-subset(my.data,subset=Site==1)
site2<-subset(my.data,subset=Site==2)
site3<-subset(my.data,subset=Site==3)
head(site3)

```


## Are there significant differences in coral cover among sites and through time?
```{r}
coverlong<-read.csv("cover_long.csv")
colnames(coverlong)[1]<-"Site"

coverlong$Site<-as.factor(coverlong$Site)
coverlong$plotnum<-as.factor(coverlong$plotnum)
coverlong$timept<-as.factor(coverlong$timept)
coverlong<-coverlong%>%
  filter(plotnum!="p24",plotnum!="p29",plotnum!="p46")%>%
  droplevels()
#remove the treated quadrats for paper analysis
mod<-aov((percent.cover)~Site*timept,data=coverlong)
plot(mod)
shapiro.test(mod$residuals)
#data do not meet assumptions for ANOVA, so an aligned ranks transformation 2-way ANOVA (ARTool package in R, Wobbrock et al. 2011) was used to determine significant differences in percent coral cover among sites and between the initial and final time points. 
summary(coverlong)

mod<-art(percent.cover~Site*timept,data=coverlong)
summary(mod)
(anova(mod))
testInteractions(artlm(mod,"Site:timept"),pairwise=c("Site","timept"))
```


## Are there significant differences in percent coral tissue loss per week among the three sites?
```{r}
my.df<-read.csv("SWG_SCTLDprogrates.csv")

kruskal.test(Avg_prweek~Site,data=my.df)

my.df$Site<-as.factor(my.df$Site)

### subset by site
tl.s1<-subset(my.df,subset=Site==1)
tl.s2<-subset(my.df,subset=Site==2)
tl.s3<-subset(my.df,subset=Site==3)

s1means<-mean(tl.s1$Avg_prweek)
s2means<-mean(tl.s2$Avg_prweek)
s3means<-mean(tl.s3$Avg_prweek)

#standard error
s1se<-my.se(tl.s1$Avg_prweek)
s2se<-my.se(tl.s2$Avg_prweek)
s3se<-my.se(tl.s3$Avg_prweek)
tl.means<-rbind(s3means,s1means,s2means)
#tl.means
tl.se<-rbind(s3se,s1se,s2se)
tl.se[is.na(tl.se)]<-0
bp<-barplot(as.matrix((tl.means)),beside=TRUE,ylim=c(0,15),ylab=strwrap("Change in percent tissue loss per week per colony with SCTLD",width=40) ,names.arg=c("Inshore","Midchannel","Offshore"),col=c("grey","light blue","blue"))
arrows(x0=bp,x1=bp,y0=(tl.means)-1.96*(tl.se),y1=(tl.means)+1.96*(tl.se),code = 3, angle = 90, len = 0.02, xpd = NA)
legend(x = 1, y=104,legend = c("Inshore","Midchannel","Offshore"), fill =c("grey","light blue","blue"),bty="n")


```

No significant differences in percent tissue loss per week per colony among sites. So we don't need to include site as a factor when looking for differences in progression rates among species.

## Differences in progression rates among species?
```{r}
aggregate(Avg_prweek~Sps,data=my.df,FUN=mean)
my.df$Sps<-as.factor(my.df$Sps)
#my.df_noplci<-subset(my.df,subset=Sps!="PCLI")
mod<-aov((Avg_prweek)~Sps,data=my.df)
plot(mod)
shapiro.test(mod$residuals)
#try transformations
mod<-aov(log(Avg_prweek)~Sps,data=my.df)
plot(mod)
shapiro.test(mod$residuals)
#resort to kruskal wallis
kruskal.test(Avg_prweek~Sps,data=my.df)

```

Significant differences among species.

```{r}
sps.avg<-aggregate(Avg_prweek~Sps,data=my.df,FUN=mean)
sps.se<-aggregate(Avg_prweek~Sps,data=my.df,FUN=my.se)
par(family="Times New Roman")
bp<-barplot(as.matrix(t(sps.avg$Avg_prweek)),ylim=c(0,25),las=1,ylab="Percent loss per week",names.arg=c(as.character(sps.avg$Sps)),las=2,col="grey")
arrows(x0=bp,x1=bp,y0=(sps.avg$Avg_prweek)-1.96*(sps.se$Avg_prweek),y1=(sps.avg$Avg_prweek)+1.96*(sps.se$Avg_prweek),code = 3, angle = 90, len = 0.02, xpd = NA)
numbers<- c("37","6","37","21","11","4","1","28","8","11","12")
text(x=bp,y=1,numbers )
#text(x = bp, y =(sps.avg$Avg_prweek)+1.96*(sps.se$Avg_prweek), labels, pos = 3)

```

## Temperature correlations?
Average progression rates and total incidence among sites were negatively associated with SST and DHW from 04 January 2019 when the disease incidence was first over 5 cases to the end of the surveys on 06 December 2019.
```{r}
envdis<-read.csv("extended_envdisFig3.csv")
colnames(envdis)<-c("dates","sst","bleachalert","dhw","t.anom","tl.means","tl.se","newInc")
summary(envdis)
#incidence and dhw
mod<-lm(log(newInc)~dhw,data=envdis[10:26,])
shapiro.test(mod$residuals)
acf(mod$residuals)
cor.test(envdis$dhw[10:26],log(envdis$newInc[10:26]))
#incidence and sst
mod<-lm(log(newInc)~sst,data=envdis[10:26,])
shapiro.test(mod$residuals)
acf(mod$residuals)
cor.test(envdis$sst[10:26],log(envdis$newInc[10:26]))
#tissue loss and dhw
mod<-lm((tl.means)~dhw,data=envdis[10:26,])
shapiro.test(mod$residuals)
acf(mod$residuals)
cor.test(envdis$dhw[10:26],log(envdis$tl.means[10:26]))
#tissue loss and sst
mod<-lm(log(tl.means)~sst,data=envdis[10:26,])
shapiro.test(mod$residuals)
acf(mod$residuals)
cor.test(envdis$sst[10:26],log(envdis$tl.means[10:26]))
log(envdis$tl.means)

```
#### Temperature Correlation figure


```{r}
colnames(envdis)<-c("dates","sst","bleachalert","dhw","t.anom","tl.means","tl.se","newInc")
tl.newI<-rbind(envdis$tl.means,envdis$newInc)
tl.newI

tl.newI.se<-rbind(envdis$tl.se,rep(0,times=length(tl.se)))
envdis$dates<-revalue(envdis$dates, c( "X05.01.18"="05-01-18","X06.01.18"="06-01-18","X06.21.18"="06-21-18","X07.16.18"="07-16-18","X08.17.18"="08-17-18","X10.30.18"="10-30-18", "X11.9.18"="11-09-18", "X11.29.18"="11-29-18","X12.13.18"="12-13-18","X1.4.19"="01-04-19","X1.18.19"="01-18-19","X2.8.19"="02-08-19","X3.4.19"="03-04-19","X3.21.19"="03-21-19","X4.11.19"="04-11-19","X5.2.19"="05-02-19","X5.16.19"="05-16-19","X5.28.19"="05-28-19","X6.13.19"="06-13-19","X7.1.19"="07-01-19","X7.22.19"="07-22-19","X8.16.19"="08-16-19","X9.17.19"="09-17-19","X10.14.19"="10-14-19","X11.12.19"="11-12-19","X12.6.19"="12-06-19"))
envdis$dates<-as.Date(envdis$dates,"%m-%d-%y")
envdis$dates<-format(envdis$dates,"%d %b %y")

#tiff("Figure3.tiff",width=180, height=120,units="mm",res=300)
par(oma = c(0, 1, 1, 3),family="Times New Roman")
bp<-barplot(as.matrix((tl.newI)),beside=TRUE,ylim=c(0,35),las=1,ylab="Disease metric", names.arg=envdis$dates,las=2,col=c("light grey","dark grey"))
legend("topleft",cex=.7,bty="n",col=c("light grey","dark grey"),fill=c("light grey","dark grey"),legend=c("Percent coral tissue loss","Incidence"))
arrows(x0=bp,x1=bp,y0=(tl.newI)-(tl.newI.se),y1=(tl.newI)+(tl.newI.se),code = 3, angle = 90, len = 0.02, xpd = NA)
par(new = T, family="Times New Roman")
plot(bp[1,],envdis$dhw,xlab=NA,ylab=NA,axes=F,ylim=c(0,10),type="b",col="blue",pch=19,lwd=2)
axis(side = 4,col="blue",line=3, lwd=2)
#mtext(side = 4, line = 3, 'Degree Heating Week')
par(new = T)
plot(bp[1,],envdis$sst,xlab=NA,ylab=NA,axes=F,ylim=c(20,34),type="b",col="red")
axis(side = 4,col="red")
mtext(side = 3, 'SST      DHW',at=86,line=1)
#dev.off()
```

## Did susceptible corals who bleached then get sctld?
```{r}
###looking at thermal stress
my.data.ts<-read.csv("SCTLD_END_Vpub_ts.csv") ## this file is the same as the original, but includes columns related to thermal stress signs
colnames(my.data.ts)[1]<-"Site"
dis.sps.ts<-my.data.ts%>%
  filter(Sps!="AAGA",Sps!="ACER",Sps!="ATEN",Sps!="EFAS",Sps!="MANG",Sps!="MMEA",Sps!="MYCE",Sps!="OCUL",Sps!="ODIF",Sps!="PAST",Sps!="PDIV",Sps!="PPOR",Sps!="SRAD")%>%
  droplevels()
dis.ts.table<-table(dis.sps.ts$tot_diseased,dis.sps.ts$tot_stressed)
chisq.test(dis.ts.table)$expected
dis.ts.table
chisq.test(dis.ts.table)
```


## Are certain species more likely to show signs of SCTLD than species? Yes.
```{r}
my.table<-table(my.data$Sps,my.data$glom)
my.table
chisq.test(my.table)$expected #does not meet assumptions for chi square
fisher.test(my.table,simulate.p.value=T)
#chisq.test(table(dis.sps$Sps,dis.sps$glom))$expected
```
Same, but for just the 11 susceptible species?
```{r}
dis.sps<-my.data%>%
  filter(Sps!="AAGA",Sps!="ACER",Sps!="ATEN",Sps!="EFAS",Sps!="MANG",Sps!="MMEA",Sps!="MYCE",Sps!="OCUL",Sps!="ODIF",Sps!="PAST",Sps!="PDIV",Sps!="PPOR",Sps!="SRAD")%>%
  droplevels()
my.table<-table(dis.sps$Sps,dis.sps$glom)
chisq.test(my.table)$expected #does not meet assumptions for chi square
fisher.test(my.table,simulate.p.value=T)

```

## Are larger corals more likely to to show signs of SCTLD?
```{r}
## let's look at all under 200cm, becuase there's just one or two outliers
mwidth.200cm<-subset(dis.sps,subset=Max_width<=200)
nrow(mwidth.200cm)/nrow(dis.sps) #99.7% under 200cm
```

```{r}
mw.means<-aggregate(Max_width~glom,data=mwidth.200cm,FUN=mean)
#mw.means
mw.se<-aggregate(Max_width~glom,data=mwidth.200cm,FUN=function(x) sd(x)/sqrt(length(x)))
#mw.se
ci.upp <- mw.means$Max_width + 1.96 * mw.se$Max_width
#ci.upp
ci.low <- mw.means$Max_width - 1.96 * mw.se$Max_width
disnames<-c("Healthy","SCTLD")
bp <- barplot(mw.means$Max_width, beside = TRUE, names = disnames,col=c("lightgrey","darkgrey"),ylim=c(0,100),ylab="Maximum Colony Width (cm)",horiz=FALSE)
arrows(y0 = ci.low, y1 = ci.upp, x0 = bp, x1 = bp, angle = 90, code = 3, length = 0.1)
```
```{r}
aggregate(Max_width~glom,data=mwidth.200cm,FUN=mean)

dis.w<-subset(mwidth.200cm,subset=glom=="SCTLD",select="Max_width")
health.w<-subset(mwidth.200cm,subset=glom=="Healthy",select="Max_width")
shapiro.test((dis.w$Max_width))
shapiro.test((health.w$Max_width))

wilcox.test((dis.w$Max_width),(health.w$Max_width)) #significant! p-value = 8.132e-12


```
We use the Wilcoxon rank sum test (Mann-Whitney) becuase the data do not meet assumptions. Compares the medians of two groups using ranks. 

```{r}
mw.means<-aggregate(Max_width~Site,data=mwidth.200cm,FUN=mean)
#mw.means
mw.se<-aggregate(Max_width~Site,data=mwidth.200cm,FUN=function(x) sd(x)/sqrt(length(x)))
#mw.se
ci.upp <- mw.means$Max_width + 1.96 * mw.se$Max_width
#ci.upp
ci.low <- mw.means$Max_width - 1.96 * mw.se$Max_width
disnames<-c("Healthy","SCTLD")
bp <- barplot(mw.means$Max_width, beside = TRUE, names = levels(mwidth.200cm$Site),ylim=c(0,100),ylab="Maximum Colony Width (cm)",horiz=FALSE)
arrows(y0 = ci.low, y1 = ci.upp, x0 = bp, x1 = bp, angle = 90, code = 3, length = 0.1)
```

## Now for difs within species, multiple comparisons and BH correction
```{r}
#table(mwidth.200cm$Sps,mwidth.200cm$glom)
#levels(as.factor(mwidth.200cm$Sps))
#mwidth.200cm$Sps<-as.factor(mwidth.200cm$Sps)
resmat<-matrix(NA,nrow=0,ncol=3)
dis.sps_np<-subset(dis.sps,subset=Sps!="PCLI") #take out pcli because only one colony
dis.sps_np$Sps<-as.factor(dis.sps_np$Sps)

#levels(dis.sps_np$Sps)
for (i in 1:length(levels(dis.sps_np$Sps))){
  #print(levels(dis.sps_np$Sps)[i])
  just.one<-subset(dis.sps_np,subset=Sps==levels(dis.sps_np$Sps)[i])
  dis.w<-subset(just.one,subset=glom=="SCTLD",select="Max_width")
  health.w<-subset(just.one,subset=glom=="Healthy",select="Max_width")
  stat<-wilcox.test(dis.w$Max_width,health.w$Max_width)
  results<-c(levels(dis.sps_np$Sps)[i], stat$statistic,stat$p.value)
  resmat<-rbind(resmat,results)
}
nomaxorpcli_all_results<-cbind(resmat,p.adjust(as.numeric(resmat[,3]),"hochberg"),p.adjust(as.numeric(resmat[,3]),"holm"),p.adjust(as.numeric(resmat[,3]),"bonferroni"))
results<-as.data.frame(nomaxorpcli_all_results)
colnames(results)<-c("sps","wstat", "p.val","benajmini_hochberg","holm", "bonferoni")
results
```
So there are significant differences for CNAT, DSTO, SSID 

```{r}
## Get Barplot
mw.means.sp<-aggregate(Max_width~Sps+glom,data=dis.sps,FUN=mean)
#mw.means.sp
mw.se.sp<-aggregate(Max_width~Sps+glom,data=dis.sps,FUN=function(x) sd(x)/sqrt(length(x)))

size.sp<-cbind(mw.means.sp,mw.se.sp$Max_width)
colnames(size.sp)<-c("sps","state","mean","se")
ggplot(size.sp,aes(x=sps,y=mean,fill=state))+
  geom_bar(stat="identity",position=position_dodge(.9))+theme(panel.background = element_blank())+
  geom_errorbar(aes(ymax=mean + se, ymin=mean-se),width=0.2,position=position_dodge(.9))+
  theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+scale_fill_manual(values=c("Healthy" = "grey52","SCTLD"="grey28"), drop = FALSE)+
  theme(legend.position="none")+scale_x_discrete(limits=c("SSID","SINT","MCAV","DSTO","CNAT","PSTR","SBOU","OANN","OFAV","DLAB","PCLI"))+
  labs(x="",y="Maximum width (cm)")+
  theme(legend.position = c(0.9, 0.7) ,legend.text=element_text(size=12),legend.title=element_blank(),legend.key.size =unit(0.5,"line"))+scale_y_continuous(breaks=seq(0,80,10),expand = c(0, 0),lim=c(0,80))
#progrates
#sps.avg
#size.sp
```

## Time of disease onset


## Now look at timing of disease
```{r}
my.data.dis<-my.data%>%
  filter(Sps!="AAGA",Sps!="ACER",Sps!="ATEN",Sps!="EFAS",Sps!="MANG",Sps!="MMEA",Sps!="MYCE",Sps!="OCUL",Sps!="ODIF",Sps!="PAST",Sps!="PDIV",Sps!="PPOR",Sps!="SRAD")%>%
  droplevels()
head(my.data.dis)
timedifs<-read.csv("timedif_surveys.csv") #survey dates, number of days since last date, and running sum of days; calculated and formatted in excel
my.data.disonly<-my.data.dis%>%
  filter(total_bin>0)
timedifs
```

need the new_inf_corals function that returnts the number of newly infected corals
```{r}
new_inf_corals<-function(df,steps,x){
  ## newly_I
  newly_I<-matrix(0, nrow = nrow(df), ncol = steps) #blank matrix for storing newly infected corals
  #newly_I
  for (i in 1:steps){ #for each survey time point
    col<-x+i #start with first tp after init tp
    prev<-x+i-1
    for (j in 1:(nrow(df))){ #for each row in the df
      #print (df[j,col])
      if ( df[j,col]=="SCTLD" ){ #if it's disease
        #print( "found one")
        if (df[j,prev]!="SCTLD"){ # and if it wasnt diseased before
          newly_I[j,i]<-1 #add it to newly infected
        }
      }
    }
  }
  return(newly_I)
}
```

```{r}
head(my.data.disonly)
```


```{r}
colnames(my.data.disonly)
wheninfected<-new_inf_corals(my.data.disonly,21,12) #df, steps, x
wheninfected
dateinf<-c()
for (i in 1:nrow(my.data.disonly)){ #for ech colony
  for (j in 1:ncol(wheninfected)){ #for each date in wheninfected
    if(wheninfected[i,j]==1){ #if a colony is infected
      dateinf[i]<-j #the date infected is J
    }
    if(my.data.disonly$total[i]==0){
      dateinf[i]<-"Healthy"
    }
  }
  
}
nrow(wheninfected)
nrow(my.data.disonly)
length(dateinf)
median(dateinf)
my.data.disonly$dateinf<-dateinf #starting in october
my.data.disonly$dateinf<-as.numeric(my.data.disonly$dateinf)

table(my.data.disonly$Site,my.data.disonly$dateinf)
site1_initdis<-3
site2_initdis<-1
site3_initdis<-7

dateinfbysite<-data.frame()
dateinfbysite

for (i in 1:nrow(my.data.disonly)){ #for each colony
  if(my.data.disonly$Site[i]==1){ #if in site 1
    dateinfbysite[i,1]<-site1_initdis #make a new column that tracks date that infection started at each site
  }
  if(my.data.disonly$Site[i]==2){
    dateinfbysite[i,1]<-site2_initdis
  }
  if(my.data.disonly$Site[i]==3){
    dateinfbysite[i,1]<-site3_initdis
  }
}
dateinfbysite
my.data.disonly$datesiteinf<-dateinfbysite[,1] #add to major dataset
my.data.disonly$dateinfbysite<-as.numeric(my.data.disonly$dateinf)-my.data.disonly$datesiteinf #subtract the infection start date from the date each colony was infected
timedifs # came from somewhere else... saved csv
dateinf_days<-c()
datesiteinf_days<-c()
for (i in 1:nrow(my.data.disonly)){ #for each colony
  dateinf_days[i]<-sum(timedifs$days_since_last[1:my.data.disonly$dateinf[i]])
  datesiteinf_days[i]<-sum(timedifs$days_since_last[1:my.data.disonly$datesiteinf[i]])
}

#timedifs
#dateinf_days
#my.data.disonly$dateinf
#datesiteinf_days
my.data.disonly$dateinfbysite_days<-dateinf_days-datesiteinf_days

boxplot(my.data.disonly$dateinfbysite_days~my.data.disonly$Sps,na.rm=TRUE,ylab="Survey Number after")

my.data.disonly$dateinfbysite_weeks<-my.data.disonly$dateinfbysite_days/7
boxplot(my.data.disonly$dateinfbysite_weeks~my.data.disonly$Sps,na.rm=TRUE,ylab="Survey Number after")


infdate.sps.avg<-aggregate(dateinfbysite_weeks~Sps,data=my.data.disonly,FUN=mean)
infdate.sps.se<-aggregate(dateinfbysite_weeks~Sps,data=my.data.disonly,FUN=my.se)
infdate.sps.se[is.na(infdate.sps.se)]<-0
par(family="Times New Roman")
bp<-barplot(as.matrix(t(infdate.sps.avg$dateinfbysite_weeks)),ylim=c(0,60),las=1,ylab="When infected (Weeks after first infection at site)",names.arg=c(as.character(infdate.sps.avg$Sps)),las=2,col="grey")
arrows(x0=bp,x1=bp,y0=(infdate.sps.avg$dateinfbysite_weeks)-1.96*(infdate.sps.se$dateinfbysite_weeks),y1=(infdate.sps.avg$dateinfbysite_weeks)+1.96*(infdate.sps.se$dateinfbysite_weeks),code = 3, angle = 90, len = 0.02, xpd = NA)

infdate.sps.se[is.na(infdate.sps.se)]<-0

inftime<-ggplot(infdate.sps.avg,aes(Sps,y=dateinfbysite_weeks))+
  geom_bar(stat="identity",fill="grey45")+theme(panel.background = element_blank())+
  geom_errorbar(aes(ymax=infdate.sps.avg$dateinfbysite_weeks +infdate.sps.se$dateinfbysite_weeks, ymin=infdate.sps.avg$dateinfbysite_weeks-infdate.sps.se$dateinfbysite_weeks),width=0.2)+
  theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  theme(legend.position="none")+scale_x_discrete(limits=c("SSID","SINT","MCAV","DSTO","CNAT","PSTR","SBOU","OANN","OFAV","DLAB","PCLI"))+
  labs(x="",y="")+
  ylab(expression(paste("Time of disease onset (weeks \n after first diseased signs  observed)")))+
  theme(legend.position = c(0.9, 0.7) ,legend.text=element_text(size=10),legend.title=element_blank(),legend.key.size =unit(0.5,"line"))+scale_y_continuous(breaks=seq(0,50,5),expand = c(0, 0),lim=c(0,50))
inftime
median(infdate.sps.avg$dateinfbysite_weeks)

```


## Reef level indicators of susceptibility

### Species Diversity Analyses
Goal: Visualize relationships between total disease prevalence per plot and it's relationships with coral diversity and density metrics
- Shannon Diversity
- Species Richness
- Species Density
- Colony Density

#### Diversity Metrics
```{r}
str(my.data)
my.data$Plot<-as.factor(my.data$Plot)
### need a df where each row is a plot and columns are species
sps.list<-(table(my.data$Plot,my.data$Sps))
sps.list
sps.df<-rbind(sps.list[1,],sps.list[2,],sps.list[3,],sps.list[4,],sps.list[5,],sps.list[6,])
row.names(sps.df)<-c("p23","p25","p27","p28","p45","p47")

### calculate metrics
sh.div<-diversity(sps.df,index="shannon")
sp.rich<-specnumber(sps.df)
mean(sp.rich)
my.se(sp.rich)
evenness.J<-sh.div/specnumber(sps.df)
sps.density<-sp.rich/100 #divide by area of plot 100m^2
col.density<-rowSums(sps.list)/100
ncorals<-rowSums(sps.list)
sp.info<-cbind(sh.div,sp.rich,sps.density,evenness.J,col.density,ncorals)
sp.info
plotprev<-aggregate(total_bin~Plot,data=my.data,FUN=function(x) sum(x)/length(x))
plotprev
avgplotsiteprev<-(plotprev[seq(from = 1, to = NROW(plotprev), by = 2),2]+plotprev[seq(from = 2,to = NROW(plotprev), by = 2),2])/2

#make dataframe for plot-level data
colnames(plotprev)<-c("Plot","totprev")
plot.df<-cbind(plotprev,sp.info)
sitenum<-c(1,1,2,2,3,3)
plot.df<-cbind(plot.df,sitenum)
plot.df
#add maxwidth
plot.df$avg.max_width<-aggregate(Max_width~Plot,FUN=mean,data=my.data)
#add cover
str(coverlong) #from percent cover analysis above
cover.means<-aggregate(percent.cover~plotnum+timept,data=coverlong,FUN=mean)
cover<-cover.means[1:6,3]
plot.df$initcover<-cover
plot.df$avg.max_width
```


```{r}
plot.data<-plot.df
plot.data$avgmaxwidth<-plot.data$avg.max_width$Max_width

div.mod<-lm(totprev~sh.div,data=plot.data)
div.plot<-ggplot(div.mod$model, aes_string(x = names(div.mod$model)[2], y = names(div.mod$model)[1])) + 
  geom_point(aes(color=c("Midchannel","Midchannel","Offshore","Offshore","Inshore","Inshore"))) +
  stat_smooth(method = "lm", col = "red",se=TRUE,size=.5,alpha=0.1)+
  theme(panel.background = element_blank())+ theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(legend.position="None")+
  labs(x="Shannon diversity",y="Total prevalence")+scale_color_manual(values=c("grey","light blue","blue"))+scale_y_continuous(breaks=seq(0,.2,.05),expand = c(0, 0),limits=c(-0.01,.25))
div.plot

dens.mod<-lm(totprev~col.density,data=plot.data)
dens.plot<-ggplot(dens.mod$model, aes_string(x = names(dens.mod$model)[2], y = names(dens.mod$model)[1])) + 
  geom_point(aes(color=c("Midchannel","Midchannel","Offshore","Offshore","Inshore","Inshore"))) +
  stat_smooth(method = "lm", col = "red",se=TRUE,size=.5,alpha=0.1,linetype=2)+
  theme(panel.background = element_blank())+ theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(legend.position="None")+
  labs(x="Colony density",y="Total prevalence",color="Site")+scale_color_manual(values=c("grey","light blue","blue"))+scale_y_continuous(breaks=seq(-.1,.3,.05),limits=c(-0.1,.3))
dens.plot
div.plot

cov.mod<-lm(totprev~initcover,data=plot.data)
cov.plot<-ggplot(cov.mod$model, aes_string(x = names(cov.mod$model)[2], y = names(cov.mod$model)[1])) + 
  geom_point(aes(color=c("Midchannel","Midchannel","Offshore","Offshore","Inshore","Inshore"))) +
  stat_smooth(method = "lm", col = "red",se=TRUE,size=.5,alpha=0.1)+
  theme(panel.background = element_blank())+ theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(legend.position="None")+
  labs(x="Initial percent coral cover",y="Total prevalence",color="Site")+scale_color_manual(values=c("grey","light blue","blue"))+scale_y_continuous(breaks=seq(0,.3,.05),expand = c(0, 0),limits=c(-0.01,.3))
cov.plot
summary(cov.mod)

size.mod<-lm(totprev~avgmaxwidth,data=plot.data)
size.plot<-ggplot(size.mod$model, aes_string(x = names(size.mod$model)[2], y = names(size.mod$model)[1])) + 
  geom_point(aes(color=c("Midchannel","Midchannel","Offshore","Offshore","Inshore","Inshore"))) +
  stat_smooth(method = "lm", col = "red",se=TRUE,size=.5,alpha=0.1)+
  theme(panel.background = element_blank())+ theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+theme(legend.position="None")+
  labs(x="Average colony maximum width",y="Total prevalence",color="Site")+scale_color_manual(values=c("grey","light blue","blue"))+scale_y_continuous(breaks=seq(0,.2,.05),expand = c(0, 0),limits=c(0,.25))
size.plot

```

```{r}
dens.plot+cov.plot+div.plot+size.plot
```

#### What species indicate a reef may be susceptible?: PCOA

```{r}
library(viridis)
library(factoextra)
library(FactoMineR)
library(corrplot)
```


```{r}
#get the number of each susceptible species per quadrat
num_dsto<-table(my.data$Plot,my.data$Sps=="DSTO")
num_mcav<-table(my.data$Plot,my.data$Sps=="MCAV")
num_pstri<-table(my.data$Plot,my.data$Sps=="PSTR")
num_cnats<-table(my.data$Plot,my.data$Sps=="CNAT")
num_dlab<-table(my.data$Plot,my.data$Sps=="DLAB")
num_ofav<-table(my.data$Plot,my.data$Sps=="OFAV")
num_oann<-table(my.data$Plot,my.data$Sps=="OANN")
num_sbou<-table(my.data$Plot,my.data$Sps=="SBOU")
num_sint<-table(my.data$Plot,my.data$Sps=="SINT")
num_ssid<-table(my.data$Plot,my.data$Sps=="SSID")


Pstr<-num_pstri[,2]
Cnat<-num_cnats[,2]
Dsto<-num_dsto[,2]
Mcav<-num_mcav[,2]
Dlab<-num_dlab[,2]
Ofav<-num_ofav[,2]
Oann<-num_oann[,2]
Sbou<-num_sbou[,2]
Sint<-num_sint[,2]
Ssid<-num_ssid[,2]
df<-cbind(Pstr,Cnat,Dsto,Mcav,Dlab,Ofav,Oann,Sbou,Sint,Ssid)
df
norm.df<-df/rowSums(df) #normalized (by total number s sps) abundances of susceptible species at the 6 plots. 

norm.df
```

```{r}
s.sps.pca<-PCA(norm.df) #performs principle component analysis on normalized species counts dataframe
get_eigenvalue(s.sps.pca) #eigenvalues measures the amount of variation in each principle component
#scree plot
fviz_eig(s.sps.pca) #visualizes hte amount of variation explained by each pc
#extract results for variables
var<-get_pca_var(s.sps.pca)
var$contrib #contrib is the contribution in percentage of the variables to the principle components
```


```{r}
corrplot(var$contrib,is.corr=FALSE)
fviz_cos2(s.sps.pca,choice="var",axes=1:2)

```
So Dlab less important for explaining most of the variation among quadrats

```{r}
fviz_contrib(s.sps.pca,choice="var",axes=1)
fviz_contrib(s.sps.pca,choice="var",axes=2)
```

```{r}
fviz_pca_var(s.sps.pca,alpha.var="contrib",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"))
```

```{r}
dimdesc(s.sps.pca,axes=c(1,2),proba=0.05)
#This function is designed to point out the variables and the categories that are the most characteristic according to each dimension obtained by a Factor Analysis. 

```

```{r}
var<-get_pca_var(s.sps.pca)
var$contrib
km <- kmeans(var$coord, centers = 2, nstart = 25)
grp <- as.factor(km$cluster)

nice.biplot<-fviz_pca_biplot(s.sps.pca, 
       
                # Fill individuals by groups
                #geom.ind = "point",
                pointshape = 21,
                pointsize = 2.5,
                mean.point=FALSE,
                fill.ind = as.factor(c("Midchannel","Midchannel","Offshore","Offshore","Nearshore","Nearshore")),
                col.ind = "black",
                # Color variable by groups
                col.var = grp,
                alpha.var ="contrib",
                
                legend.title = list(fill = "Site", alpha="Contribution",color="Cluster"),
                repel=TRUE,
                geom.ind=c("point","text"),
                axes.linetype="dashed") +
  ggpubr::fill_palette(c("blue","grey","light blue"))+
  ggpubr::color_palette(c("dark grey","black"))
nice.biplot
```


```{r}
right.biplot<-ggpubr::ggpar(nice.biplot,
              title="",
              ggtheme=theme_classic(),legend="right",ylab="PC 2",xlab="PC 1")
good.biplot<-right.biplot+
  theme(panel.background = element_blank())+ theme(text = element_text(family = "Times New Roman"))
```

```{r}
#tiff("Figure5.tiff",width=180, height=210,units="mm",res=300)
patch<-(dens.plot/div.plot) | (cov.plot/size.plot)
(patch | good.biplot) + plot_annotation(tag_levels="a")+ plot_layout(widths=c(1.5,1.5,3))+theme(text = element_text(family = "Times New Roman",size=12))
#dev.off()
```

```{r}
#tiff("Figure5.tiff",width=180, height=300,units="mm",res=300)
#patch<-(dens.plot/div.plot) | (cov.plot/size.plot)
(patch / good.biplot) + plot_annotation(tag_levels="A")+ plot_layout(widths=c(1.5,1.5,3))+theme(text = element_text(family = "Times New Roman",size=12))
#dev.off()
```


### Making Figure 4

Prevalence, Panel A
```{r}
prevplot<-ggplot(dis.sps,aes(x=Sps,fill=glom))+geom_bar(stat="count")+theme(panel.background = element_blank())+ theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x="",y="")+
  theme(legend.position="none")+scale_x_discrete(limits=c("SSID","SINT","MCAV","DSTO","CNAT","PSTR","SBOU","OANN","OFAV","DLAB","PCLI"))+labs(x="",y="Number of Colonies")+scale_fill_manual(values=c("Healthy" = "grey85","SCTLD"="grey45"), drop = FALSE)+theme(legend.position = c(0.9, 0.7) ,legend.text=element_text(size=12),legend.title=element_blank(),legend.key.size =unit(0.5,"line"))+scale_y_continuous(breaks=seq(0,650,50),expand = c(0, 0),lim=c(0,650))+theme(axis.text.x = element_text(angle = 45,hjust=1))
prevplot
```

Percent tissue loss per week, panel B
```{r}
#sps.avg
#sps.se
prog.fig<-cbind(sps.avg,sps.se)
colnames(prog.fig)<-c("sps","mean","sps2","se")

progrates<-ggplot(prog.fig,aes(sps,y=mean))+
  geom_bar(stat="identity",fill="grey45")+theme(panel.background = element_blank())+
  geom_errorbar(aes(ymax=mean + se, ymin=mean-se),width=0.2)+
  theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  theme(legend.position="none")+scale_x_discrete(limits=c("SSID","SINT","MCAV","DSTO","CNAT","PSTR","SBOU","OANN","OFAV","DLAB","PCLI"))+
  labs(x="",y="Percent tissue loss per week")+
  theme(legend.position = c(0.9, 0.7) ,legend.text=element_text(size=10),legend.title=element_blank(),legend.key.size =unit(0.5,"line"))+scale_y_continuous(breaks=seq(0,20,2),expand = c(0, 0),lim=c(0,20))+theme(axis.text.x = element_text(angle = 45,hjust=1))
progrates

```

Infection timing, panel C
```{r}
#From above
infdate.sps.avg
infdate.sps.se
inftime<-ggplot(infdate.sps.avg,aes(Sps,y=dateinfbysite_weeks))+
  geom_bar(stat="identity",fill="grey45")+theme(panel.background = element_blank())+
  geom_errorbar(aes(ymax=infdate.sps.avg$dateinfbysite_weeks +infdate.sps.se$dateinfbysite_weeks, ymin=infdate.sps.avg$dateinfbysite_weeks-infdate.sps.se$dateinfbysite_weeks),width=0.2)+
  theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  theme(legend.position="none")+scale_x_discrete(limits=c("SSID","SINT","MCAV","DSTO","CNAT","PSTR","SBOU","OANN","OFAV","DLAB","PCLI"))+
  labs(x="",y="")+
  ylab(expression(paste("Time of disease onset (weeks \n after first diseased signs  observed)")))+
  theme(legend.position = c(0.9, 0.7) ,legend.text=element_text(size=10),legend.title=element_blank(),legend.key.size =unit(0.5,"line"))+scale_y_continuous(breaks=seq(0,50,5),expand = c(0, 0),lim=c(0,50))+theme(axis.text.x = element_text(angle = 45,hjust=1))
inftime
median(infdate.sps.avg$dateinfbysite_weeks)
infdate.sps.avg
infdate.sps.se
```

Size and disease panel D
```{r}
#size.sp<-read.csv("size_sp.csv")
size.sp.plot<-ggplot(size.sp,aes(x=sps,y=mean,fill=state))+
  geom_bar(stat="identity",position=position_dodge(.9))+theme(panel.background = element_blank())+
  geom_errorbar(aes(ymax=mean + se, ymin=mean-se),width=0.2,position=position_dodge(.9))+
  theme(text = element_text(family = "Times New Roman"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+scale_fill_manual(values=c("Healthy" = "grey85","SCTLD"="grey45"), drop = FALSE)+
  theme(legend.position="none")+scale_x_discrete(limits=c("SSID","SINT","MCAV","DSTO","CNAT","PSTR","SBOU","OANN","OFAV","DLAB","PCLI"))+
  labs(x="",y="Maximum width (cm)")+
  theme(legend.position = "none" ,legend.text=element_text(size=10),legend.title=element_blank(),legend.key.size =unit(0.5,"line"))+scale_y_continuous(breaks=seq(0,80,10),expand = c(0, 0),lim=c(0,80))+
  geom_signif(comparisons=list(c("CNAT","CNAT"), c("SSID","SSID"),c("DSTO","DSTO")), annotations="***",
              y_position = c(64,49,16), tip_length = 0, vjust=0.4)+theme(axis.text.x = element_text(angle = 45,hjust=1))

size.sp.plot
```

```{r, fig.height=12,fig.width=10}
#tiff("Figure4.tiff",width=85, height=260,units="mm",res=300)
(prevplot/progrates/inftime/size.sp.plot)+ plot_annotation(tag_levels="A")
#dev.off()
```
