---
title: "Ripleys K Analyses"
output: html_notebook
---
## This notebook includes:

- Data-wrangling for spatial dynamics, using "SCTLD_END_exta.csv"

- Functions for running and plotting Ripley's K analyses

- Figure 7 in paper

```{r, message=FALSE, warning=FALSE}
library(zoo) # has rollmeans function for calculating AUC
library(igraph)
library(spdep)
library(spatstat)
library(abind)
library(sciplot)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(ggpubr)
library(vegan)
library(nlme)
library(car)
library(patchwork)
library(viridis)
getwd()
```

```{r}
my.data<-read.csv("SCTLD_END_exta.csv") ## this file is shortened and doesnt included the treatment sites

#move to a long format so that every row is now an observation of a colony at a single timepoint
data_long <- gather(my.data, key=timept, value=state, X5.1.18:X12.6.19,factor_key = TRUE)
data_long<-data_long[,-c(8,9)] #drop column 8:10, which is glom and not needed anymore
#head(data_long)
#make the health states factors
data_long$state<-factor(data_long$state,levels=c("Healthy","SCTLD","Dead","Unknown"))
#revalue the timepoints
data_long$timept<-revalue(data_long$timept, c( "X5.1.18"="05-10-18","X6.1.18"="06-01-18","X6.21.18"="06-21-18","X7.16.18"="07-16-18","X8.17.18"="08-17-18","X10.30.18"="10-30-18", "X11.9.18"="11-09-18", "X11.29.18"="11-29-18","X12.13.18"="12-13-18","X1.4.19"="01-04-19","X1.18.19"="01-18-19","X2.8.19"="02-08-19","X3.4.19"="03-04-19","X3.21.19"="03-21-19","X4.11.19"="04-11-19","X5.2.19"="05-02-19","X5.16.19"="05-16-19","X5.28.19"="05-28-19","X6.13.19"="06-13-19","X7.1.19"="07-01-19","X7.22.19"="07-22-19","X8.16.19"="08-16-19","X9.17.19"="09-17-19","X10.14.19"="10-14-19","X11.12.19"="11-12-19","X12.6.19"="12-06-19"))
head(data_long)
summary(data_long$Plot)

##keep only susceptible species
data_filtered<-data_long%>%
  filter(Sps!="AAGA",Sps!="ACER",Sps!="ATEN",Sps!="EFAS",Sps!="MANG",Sps!="MMEA",Sps!="MYCE",Sps!="OCUL",Sps!="ODIF",Sps!="PAST",Sps!="PDIV",Sps!="PPOR",Sps!="SRAD")%>%
  droplevels()

Plot45 <-data_filtered %>%
  dplyr::filter(coords_x!="NA",Plot==45)

Plot47 <- data_filtered %>%
  dplyr::filter(coords_x!="NA",Plot==47)

```
## Ripley's K
Into timepoints of disease colonies per plot 
```{r}

p47<-as.data.frame(Plot47)

p45<-as.data.frame(Plot45)
#str(p47)
#now just disease
p47dis<-as.data.frame(filter(Plot47,state=="SCTLD"))

p45dis<-as.data.frame(filter(Plot45,state=="SCTLD"))


#split 47 disease tps
#start on March 4 because that's when disease cases >5 for each
p47dis_tp1<-(subset(p47dis,timept=="03-04-19",select=c(coords_x,coords_y)))

p47dis_tp2<-(subset(p47dis,timept=="03-21-19",select=c(coords_x,coords_y)))

p47dis_tp3<-(subset(p47dis,timept=="04-11-19",select=c(coords_x,coords_y)))

p47dis_tp4<-(subset(p47dis,timept=="05-02-19",select=c(coords_x,coords_y)))

p47dis_tp5<-(subset(p47dis,timept=="05-16-19",select=c(coords_x,coords_y)))

p47dis_tp6<-(subset(p47dis,timept=="05-28-19",select=c(coords_x,coords_y)))

p47dis_tp7<-(subset(p47dis,timept=="06-13-19",select=c(coords_x,coords_y)))

p47dis_tp8<-(subset(p47dis,timept=="07-01-19",select=c(coords_x,coords_y)))
p47dis_tp9<-(subset(p47dis,timept=="07-22-19",select=c(coords_x,coords_y)))
p47dis_tp10<-(subset(p47dis,timept=="08-16-19",select=c(coords_x,coords_y)))
p47dis_tp11<-(subset(p47dis,timept=="09-17-19",select=c(coords_x,coords_y)))
p47dis_tp12<-(subset(p47dis,timept=="10-14-19",select=c(coords_x,coords_y)))
p47dis_tp13<-(subset(p47dis,timept=="11-12-19",select=c(coords_x,coords_y)))
p47dis_tp14<-(subset(p47dis,timept=="12-06-19",select=c(coords_x,coords_y)))

#split 45 disease tps
p45dis_tp1<-(subset(p45dis,timept=="03-04-19",select=c(coords_x,coords_y)))

p45dis_tp2<-(subset(p45dis,timept=="03-21-19",select=c(coords_x,coords_y)))

p45dis_tp3<-(subset(p45dis,timept=="04-11-19",select=c(coords_x,coords_y)))

p45dis_tp4<-(subset(p45dis,timept=="05-02-19",select=c(coords_x,coords_y)))

p45dis_tp5<-(subset(p45dis,timept=="05-16-19",select=c(coords_x,coords_y)))

p45dis_tp6<-(subset(p45dis,timept=="05-28-19",select=c(coords_x,coords_y)))

p45dis_tp7<-(subset(p45dis,timept=="06-13-19",select=c(coords_x,coords_y)))

p45dis_tp8<-(subset(p45dis,timept=="07-01-19",select=c(coords_x,coords_y)))
p45dis_tp9<-(subset(p45dis,timept=="07-22-19",select=c(coords_x,coords_y)))
p45dis_tp10<-(subset(p45dis,timept=="08-16-19",select=c(coords_x,coords_y)))
p45dis_tp11<-(subset(p45dis,timept=="09-17-19",select=c(coords_x,coords_y)))
p45dis_tp12<-(subset(p45dis,timept=="10-14-19",select=c(coords_x,coords_y)))
p45dis_tp13<-(subset(p45dis,timept=="11-12-19",select=c(coords_x,coords_y)))
p45dis_tp14<-(subset(p45dis,timept=="12-06-19",select=c(coords_x,coords_y)))

#get all cols at one timepoint
p47.all<-(subset(p47,timept=="06-13-19",select=c(coords_x,coords_y)))

#p47.all
p45.all<-(subset(p45,timept=="06-13-19",select=c(coords_x,coords_y)))

```

### Custom Functions for running Ripley's K analyses
```{r}
#Runs and Plots results of Ripleys K
myfunc.ppp.Kest<-function(data,w,name){
  data.ppp <- as.ppp(data,w)
  # plot(data.ppp)
  lhatK <-Kest(data.ppp)
  #dividing by pi and taking the square root, linearizes the Ripley's value
  plot(lhatK$r,sqrt(lhatK$iso/pi)-lhatK$r,cex.axis=1.5,cex.lab=1.5,xlab="r (m)",col="green",ylab="L(r)",lty=1,main=name)
  # plot(lhatK$r,lhatK$iso,cex.axis=1.5,cex.lab=1.5,xlab="r (m)",col="green",ylab="L(r)",lty=1,main=name)
  return(c(data.ppp,lhatK)) #return the lhatK and ppp for use later if needed
}

#### Function to make null distribution Plot
#allpts & dispts should be my normal dataframes
myfunc.nulldistrKest<-function(allpts,dispts,w,n,name){
  #get info for all
  all<-as.data.frame(t(allpts))
  allX<-t(all)
  
  all.spp <- unique(as.ppp(allX,w))
  #print(anyDuplicated(all.spp))
  all.lhat <-Kest(all.spp)
  
  #get info for dis
  dis<-as.data.frame(t(dispts))
  disX<-t(dis)
  dis.spp <- unique(as.ppp(disX,w))
  #print(anyDuplicated(dis.spp))
  dis.lhat <-Kest(dis.spp)
  
  plot(dis.lhat$r,sqrt(dis.lhat$iso/pi)-dis.lhat$r,cex.axis=1.5,cex.lab=1.5,xlab="r (m)",ylab="L(r)",lty=0,pch=NA,main=name)
  for (i in 1:1000){
    all.samp<-sample(all,n,replace=FALSE)# n=total number of disease points to sample from all point in the plot
    x <-t(all.samp)
    all.samp.spp <- unique(as.ppp(x,w))
    #print(anyDuplicated(all.samp.spp))
    all.samp.lhat <-Kest(all.samp.spp)
    lines(all.samp.lhat$r,sqrt(all.samp.lhat$iso/pi)-all.samp.lhat$r,xlab="r (meters)",col="gray",lty=1,lwd=1)
  }
  lines(all.lhat$r,sqrt(all.lhat$theo/pi)-all.lhat$r,lwd=1,lty=8,col="black")
  lines(dis.lhat$r,sqrt(dis.lhat$iso/pi)-dis.lhat$r,lwd=1,lty=1,col="red")
  Hmisc::minor.tick(nx=10, ny=5, tick.ratio=0.5)
}


#### Functions to get difference in L(R) between actual and null distributions ######

### Function to get top of null distribution #####
top_of_RKnull<-function(allpts,w,n){
  all<-as.data.frame(t(allpts))
  allX<-t(all)
  all.spp <- unique(as.ppp(allX,w))
  all.rk <-Kest(all.spp)
  all.lhat_adj<-sqrt(all.rk$iso/pi)-all.rk$r
  plot(all.rk$r,sqrt(all.rk$iso/pi)-all.rk$r,cex.axis=1.5,cex.lab=1.5,xlab="r (m)",ylab="L(r)",lty=1)
  for (i in 1:1000){
    all.samp<-sample(all,n,replace=FALSE)# n=total number of disease points to sample from all point in the plot
    x <-t(all.samp)
    all.samp.spp <- unique(as.ppp(x,w))
    #print(anyDuplicated(all.samp.spp))
    all.samp.rk <-Kest(all.samp.spp)
    l_adj<-sqrt(all.samp.rk$iso/pi)-all.samp.rk$r
    all.lhat_adj<-cbind(all.lhat_adj,l_adj)
    lines(all.samp.rk$r,sqrt(all.samp.rk$iso/pi)-all.samp.rk$r,xlab="r (meters)",col="gray",lty=1,lwd=1)
    
  }
  ##radius stays same each time
  linemaxs<-colSums(all.lhat_adj)
  maxline<-max(linemaxs)
  best_index<-which(linemaxs==max(colSums(all.lhat_adj)))
  null_line<-all.lhat_adj[,best_index]
  return(null_line)
}

### Function to get difference between top of null and disease line #####
lhat_diffs<-function(null_line,dispts,w,n){
  dis<-as.data.frame(t(dispts))
  disX<-t(dis)
  dis.spp <- unique(as.ppp(disX,w))
  dis.rk <-Kest(dis.spp)
  dis.lhat_adj<-sqrt(dis.rk$iso/pi)-dis.rk$r
  lhat_dif<-dis.lhat_adj-null_line
  plot(dis.rk$r,lhat_dif,type="l")
  abline(h=0)
  results<-cbind(dis.rk$r,dis.lhat_adj,null_line,lhat_dif)
  return(results)
}


### Function to get top of null distribution WITH PLOT #####
top_of_RKnull_keepforplot<-function(allpts,w,n){
  all<-as.data.frame(t(allpts))
  allX<-t(all)
  all.spp <- unique(as.ppp(allX,w))
  all.rk <-Kest(all.spp)
  all.lhat_adj<-sqrt(all.rk$iso/pi)-all.rk$r
  my.mat<-matrix(nrow=513,ncol=1000)
  plot(all.rk$r,sqrt(all.rk$iso/pi)-all.rk$r,cex.axis=1.5,cex.lab=1.5,xlab="r (m)",ylab="L(r)",lty=1)
  
  for (i in 1:1000){
    all.samp<-sample(all,n,replace=FALSE)# n=total number of disease points to sample from all point in the plot
    x <-t(all.samp)
    all.samp.spp <- unique(as.ppp(x,w))
    #print(anyDuplicated(all.samp.spp))
    all.samp.rk <-Kest(all.samp.spp)
    l_adj<-sqrt(all.samp.rk$iso/pi)-all.samp.rk$r
    all.lhat_adj<-cbind(all.lhat_adj,l_adj)
    lines(all.samp.rk$r,sqrt(all.samp.rk$iso/pi)-all.samp.rk$r,xlab="r (meters)",col="gray",lty=1,lwd=1)
    my.mat[,i]<-sqrt(all.samp.rk$iso/pi)-all.samp.rk$r
  }
  ##radius stays same each time
  linemaxs<-colSums(all.lhat_adj)
  maxline<-max(linemaxs)
  best_index<-which(linemaxs==max(colSums(all.lhat_adj)))
  null_line<-all.lhat_adj[,best_index]
  my.df<-data.frame(my.mat)
  results.df<-cbind(null_line,my.mat)
  return(results.df)
}


##### func for using prev func output for plot #######
myfunc.nullplot2<-function(null_line,dispts,allpts,w,n,name){
  #get info for all
  all<-as.data.frame(t(allpts))
  allX<-t(all)
  
  all.spp <- unique(as.ppp(allX,w))
  #print(anyDuplicated(all.spp))
  all.lhat <-Kest(all.spp)
  #get info for dis
  dis<-as.data.frame(t(dispts))
  disX<-t(dis)
  dis.spp <- unique(as.ppp(disX,w))
  #print(anyDuplicated(dis.spp))
  dis.lhat <-Kest(dis.spp)
  
  plot(dis.lhat$r,sqrt(dis.lhat$iso/pi)-dis.lhat$r,cex.axis=1.5,cex.lab=1.5,xlab="r (m)",ylab="L(r)",lty=0,pch=NA,main=name)
  for (i in 1:1000){
    y<-null_line[,i+1]
    lines(all.lhat$r[0:nrow(null_line)],y,xlab="r (meters)",col="gray",lty=1,lwd=1)
  }
  lines(all.lhat$r,sqrt(all.lhat$theo/pi)-all.lhat$r,lwd=1,lty=8,col="black")
  lines(dis.lhat$r,sqrt(dis.lhat$iso/pi)-dis.lhat$r,lwd=1,lty=1,col="red")
  Hmisc::minor.tick(nx=10, ny=5, tick.ratio=0.5)
}
```

### Creating the null distribution and making the plots in the supplemental 

```{r}
w <- owin(c(-20,20),c(-20,20))
n <- 100
allpts<-p47.all
dispts<-p47dis_tp6
name<-"Quadrat 47 5/28/19"
myfunc.nulldistrKest(allpts,dispts,w,n,name)

```

Since the null distribution is random each time, I saved the null distribution for each plot once and used the same null distribution for each time point


```{r}
p47.nl1<-read.csv("plot47RKnulldist.csv",colClasses = "numeric")

myfunc.nullplot2(p47.nl1,p47dis_tp6,p47.all ,w,n=100,"Plot 47, May 28th 2019")

p45.nl1<-read.csv("plot45RKnulldist.csv",colClasses = "numeric")
myfunc.nullplot2(p45.nl1,p45dis_tp6,p45.all ,w,n=100,"Plot 45, May 28th 2019")

```

### Exctracting R, R range, and ABC from Ripley's K analyses for each timepoint

```{r}
#Saved just the top of the null distribution for each each quadrat which is used for getting values.
p47.nl1<-read.csv("topofthenull_plot47.csv",header=T)
p45.nl1<-read.csv("topofthenull_plot45.csv",header=T)
w <- owin(c(-20,20),c(-20,20))
n <- 100
head(p45.nl1)
```

```{r}
w <- owin(c(-20,20),c(-20,20))
n <- 100
# GET DIFS BETWEEN ALL TP DIS AND TOP OF NULL
p45ld.tp1<-lhat_diffs(p45.nl1[,2],p45dis_tp1,w,n)
p45ld.tp2<-lhat_diffs(p45.nl1[,2],p45dis_tp2,w,n)
p45ld.tp3<-lhat_diffs(p45.nl1[,2],p45dis_tp3,w,n)
p45ld.tp4<-lhat_diffs(p45.nl1[,2],p45dis_tp4,w,n)
p45ld.tp5<-lhat_diffs(p45.nl1[,2],p45dis_tp5,w,n)
p45ld.tp6<-lhat_diffs(p45.nl1[,2],p45dis_tp6,w,n)
p45ld.tp7<-lhat_diffs(p45.nl1[,2],p45dis_tp7,w,n)
p45ld.tp8<-lhat_diffs(p45.nl1[,2],p45dis_tp8,w,n)
p45ld.tp9<-lhat_diffs(p45.nl1[,2],p45dis_tp9,w,n)
p45ld.tp10<-lhat_diffs(p45.nl1[,2],p45dis_tp10,w,n)
p45ld.tp11<-lhat_diffs(p45.nl1[,2],p45dis_tp11,w,n)
p45ld.tp12<-lhat_diffs(p45.nl1[,2],p45dis_tp12,w,n)
p45ld.tp13<-lhat_diffs(p45.nl1[,2],p45dis_tp13,w,n)
p45ld.tp14<-lhat_diffs(p45.nl1[,2],p45dis_tp14,w,n)

### now for plot 47

p47ld.tp1<-lhat_diffs(p47.nl1[,2],p47dis_tp1,w,n)
p47ld.tp2<-lhat_diffs(p47.nl1[,2],p47dis_tp2,w,n)
p47ld.tp3<-lhat_diffs(p47.nl1[,2],p47dis_tp3,w,n)
p47ld.tp4<-lhat_diffs(p47.nl1[,2],p47dis_tp4,w,n)
p47ld.tp5<-lhat_diffs(p47.nl1[,2],p47dis_tp5,w,n)
p47ld.tp6<-lhat_diffs(p47.nl1[,2],p47dis_tp6,w,n)
p47ld.tp7<-lhat_diffs(p47.nl1[,2],p47dis_tp7,w,n)
p47ld.tp8<-lhat_diffs(p47.nl1[,2],p47dis_tp8,w,n)
p47ld.tp9<-lhat_diffs(p47.nl1[,2],p47dis_tp9,w,n)
p47ld.tp10<-lhat_diffs(p47.nl1[,2],p47dis_tp10,w,n)
p47ld.tp11<-lhat_diffs(p47.nl1[,2],p47dis_tp11,w,n)
p47ld.tp12<-lhat_diffs(p47.nl1[,2],p47dis_tp12,w,n)
p47ld.tp13<-lhat_diffs(p47.nl1[,2],p47dis_tp13,w,n)
p47ld.tp14<-lhat_diffs(p47.nl1[,2],p47dis_tp14,w,n)

```

```{r}
p47.ldiffs<-(cbind(p47ld.tp1[,1],p47ld.tp1[,4],p47ld.tp2[,4],p47ld.tp3[,4],p47ld.tp4[,4],p47ld.tp5[,4],p47ld.tp6[,4],p47ld.tp7[,4],p47ld.tp8[,4],p47ld.tp9[,4],p47ld.tp10[,4],p47ld.tp11[,4],p47ld.tp12[,4],p47ld.tp13[,4],p47ld.tp14[,4]))
p47.ldiffs<-data.frame(p47.ldiffs)
colnames(p47.ldiffs)<-c("r","03-04-19","03-21-19","04-11-19","05-02-19","05-16-19","05-28-19","06-13-19","07-01-19","07-22-19","08-16-19","09-17-19","10-14-19","11-12-19","12-06-19")

#now plot 45
p45.ldiffs<-(cbind(p45ld.tp1[,1],p45ld.tp1[,4],p45ld.tp2[,4],p45ld.tp3[,4],p45ld.tp4[,4],p45ld.tp5[,4],p45ld.tp6[,4],p45ld.tp7[,4],p45ld.tp8[,4],p45ld.tp9[,4],p45ld.tp10[,4],p45ld.tp11[,4],p45ld.tp12[,4],p45ld.tp13[,4],p45ld.tp14[,4]))
p45.ldiffs<-data.frame(p45.ldiffs)
colnames(p45.ldiffs)<-c("r","03-04-19","03-21-19","04-11-19","05-02-19","05-16-19","05-28-19","06-13-19","07-01-19","07-22-19","08-16-19","09-17-19","10-14-19","11-12-19","12-06-19")
#nrow(p45.ldiffs)
```

#### Find R, the peak clustering radius

For Quadrat/Plot 45, there are two peak clustering radii
```{r}
#look over the 25 point range around the max peak to get the best value of it. 
maxdiff.low<-matrix(nrow=25,ncol=14)
radius.low<-matrix(nrow=25,ncol=14)

maxdiff.high<-matrix(nrow=25,ncol=14)
radius.high<-matrix(nrow=25,ncol=14)
###lower clusters
p45.ldiffs.copy<-p45.ldiffs[1:205,]
for (i in 2:ncol(p45.ldiffs.copy)){
  p45.ldiffs.copy<-p45.ldiffs[1:205,]
  for (j in 1:25){
    maxdiff.low[j,i-1]<-max(p45.ldiffs.copy[,i]) 
    index<-(which(p45.ldiffs.copy[,i]==max(p45.ldiffs.copy[,i])))
    #print(index[1])
    crad<-p45.ldiffs.copy$r[index[1]]
    #print(crad)
    radius.low[j,i-1]<-crad
    p45.ldiffs.copy<-p45.ldiffs.copy[-index,]
  }
}
p45.ldiffs.copy<-p45.ldiffs[205:513,]
### upper clusters
for (i in 2:ncol(p45.ldiffs.copy)){
  p45.ldiffs.copy<-p45.ldiffs[205:513,]
  for (j in 1:25){
    maxdiff.high[j,i-1]<-max(p45.ldiffs.copy[,i]) 
    index<-(which(p45.ldiffs.copy[,i]==max(p45.ldiffs.copy[,i])))
    #print(index[1])
    crad<-p45.ldiffs.copy$r[index[1]]
    #print(crad)
    radius.high[j,i-1]<-crad
    p45.ldiffs.copy<-p45.ldiffs.copy[-index,]
  }
}

radmat.p45.low<-radius.low
maxdiffmat.p45.low<-maxdiff.low
radmat.p45.high<-radius.high
maxdiffmat.p45.high<-maxdiff.high
maxdiffmat.p45.low[,3]
colMeans(radmat.p45.high)
colMeans(radmat.p45.low)


```

For quadrat/plot 47, there is just one peak clustering radius
```{r}
maxdiff<-matrix(nrow=50,ncol=14)
radius<-matrix(nrow=50,ncol=14)
p47.ldiffs.copy<-p47.ldiffs
#p47.ldiffs.copy
p47.ldiffs.copy<-p47.ldiffs
for (i in 2:ncol(p47.ldiffs.copy)){
  p47.ldiffs.copy<-p47.ldiffs
  for (j in 1:50){
    maxdiff[j,i-1]<-max(p47.ldiffs.copy[,i])
    index<-(which(p47.ldiffs.copy[,i]==max(p47.ldiffs.copy[,i])))
    #print(index[1])
    crad<-p47.ldiffs.copy$r[index[1]]
    #print(crad)
    radius[j,i-1]<-crad
    p47.ldiffs.copy<-p47.ldiffs.copy[-index,]
  }
}

radmat.p47<-radius
maxdiffmat.p47<-maxdiff
colMeans(radmat.p47)
```

```{r}
upR45<-colMeans(radmat.p45.high)
lowR45<-colMeans(radmat.p45.low)
R47<-colMeans(radmat.p47)
datepts<-c(1,2,3,4,5,6,7,8,9,10,11,12,13,14)

plot(datepts,upR45,ylim=c(0,10),t="l",col="green")
lines(datepts,lowR45,col="blue")
lines(datepts,R47,col="black",pch=2)
```

```{r}
R.results<-cbind(c("03-04-19","03-21-19","04-11-19","05-02-19","05-16-19","05-28-19","06-13-19","07-01-19","07-22-19","08-16-19","09-17-19","10-14-19","11-12-19","12-06-19"),
upR45,lowR45,R47)

```

#### Range of significant clustering

plot 47 first
```{r}
top_range<-c()
bottom_range<-c()
#so we want to get the range of significant R
for (j in 2:ncol(p47.ldiffs)){
  #print(j)
  #print(R47[j-1])
  #the top half
  top_start<-which(p47.ldiffs[,1]>=R47[j-1])[1]
  #print(top_start)
  top_ranges<-c()
  for(i in top_start:nrow(p47.ldiffs)-1){
   if(p47.ldiffs[i,j]>=0 & p47.ldiffs[i+1,j]<=0){
     #print ("intercept")
     top_ranges<-c(top_ranges,p47.ldiffs[i,1])
    }
  }
  if(is.null(top_ranges)){
    #print("inf")
    top_ranges<-10
  }
  top_range<-c(top_range,top_ranges[1])
  #print(top_range)

  #the bottom half
  bottom_start<-which(p47.ldiffs[,1]<=R47[j-1])
  bottom_start<-bottom_start[length(bottom_start)]
  bottom_start
  bottom_ranges<-c()
  for(i in 1:bottom_start){
    if(p47.ldiffs[i,j]<=0 & p47.ldiffs[i+1,j]>=0){
     #print ("intercept")
      bottom_ranges<-c(bottom_ranges,p47.ldiffs[i,1])
    }
  }
  bottom_range<-c(bottom_range, bottom_ranges[length(bottom_ranges)])
  #print(bottom_range)
}
length(bottom_range)
length(R47)
length(top_range)
top_range
R47
R47_withrange<-cbind(bottom_range,R47,top_range)
```

now plot 45, higher R
```{r}
top_range<-c()
bottom_range<-c()
#so we want to get the range of significant R
for (j in 2:ncol(p45.ldiffs)){
  #print(j)
  #print(R47[j-1])
  #the top half
  top_start<-which(p45.ldiffs[,1]>=upR45[j-1])[1]
  #print(top_start)
  top_ranges<-c()
  for(i in top_start:nrow(p45.ldiffs)-1){
   if(p45.ldiffs[i,j]>=0 & p45.ldiffs[i+1,j]<=0){
     #print ("intercept")
     top_ranges<-c(top_ranges,p45.ldiffs[i,1])
    }
  }
  if(is.null(top_ranges)){
    #print("inf")
    top_ranges<-10
  }
  top_range<-c(top_range,top_ranges[1])
  #print(top_range)

  #the bottom half
  bottom_start<-which(p45.ldiffs[,1]<=upR45[j-1])
  bottom_start<-bottom_start[length(bottom_start)]
  bottom_start
  bottom_ranges<-c()
  for(i in 1:bottom_start){
    if(p45.ldiffs[i,j]<=0 & p45.ldiffs[i+1,j]>=0){
     #print ("intercept")
      bottom_ranges<-c(bottom_ranges,p45.ldiffs[i,1])
    }
  }
  bottom_range<-c(bottom_range, bottom_ranges[length(bottom_ranges)])
  #print(bottom_range)
}
length(bottom_range)
length(upR45)
length(top_range)

upR45_withrange<-cbind(bottom_range,upR45,top_range)
upR45_withrange
#row 5 may not be good 
```
Plot 45, lower R
```{r}
top_range<-c()
bottom_range<-c()
#so we want to get the range of significant R
for (j in 2:ncol(p45.ldiffs)){
  #print(j)
  #print(R47[j-1])
  #the top half
  top_start<-which(p45.ldiffs[,1]>=lowR45[j-1])[1]
  #print(top_start)
  top_ranges<-c()
  for(i in top_start:nrow(p45.ldiffs)-1){
   if(p45.ldiffs[i,j]>=0 & p45.ldiffs[i+1,j]<=0){
     #print ("intercept")
     top_ranges<-c(top_ranges,p45.ldiffs[i,1])
    }
  }
  if(is.null(top_ranges)){
    #print("inf")
    top_ranges<-10
  }
  top_range<-c(top_range,top_ranges[1])
  #print(top_range)

  #the bottom half
  bottom_start<-which(p45.ldiffs[,1]<=lowR45[j-1])
  bottom_start<-bottom_start[length(bottom_start)]
  bottom_start
  bottom_ranges<-c()
  for(i in 1:bottom_start){
    if(p45.ldiffs[i,j]<=0 & p45.ldiffs[i+1,j]>=0){
     #print ("intercept")
      bottom_ranges<-c(bottom_ranges,p45.ldiffs[i,1])
    }
  }
  bottom_range<-c(bottom_range, bottom_ranges[length(bottom_ranges)])
  #print(bottom_range)
}
length(bottom_range)
length(lowR45)
length(top_range)

lowR45_withrange<-cbind(bottom_range,lowR45,top_range)
lowR45_withrange
#row 5 may not be good 
```

### Area under/between the curve

```{r}
p47.ldiffsc<-p47.ldiffs
p47.ldiffsc[p47.ldiffsc<0]<-0

p47.tp1AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,2],2))
p47.tp2AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,3],2))
p47.tp3AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,4],2))
p47.tp4AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,5],2))
p47.tp5AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,6],2))
p47.tp6AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,7],2))
p47.tp7AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,8],2))
p47.tp8AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,9],2))
p47.tp9AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,10],2))
p47.tp10AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,11],2))
p47.tp11AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,12],2))
p47.tp12AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,13],2))
p47.tp13AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,14],2))
p47.tp14AUC <- sum(diff(p47.ldiffsc[,1])*rollmean(p47.ldiffsc[,15],2))

p47.AUC<-rbind(p47.tp1AUC,p47.tp2AUC,p47.tp3AUC,p47.tp4AUC,p47.tp5AUC,p47.tp6AUC,p47.tp7AUC,p47.tp8AUC,p47.tp9AUC,p47.tp10AUC,p47.tp11AUC,p47.tp12AUC,p47.tp13AUC,p47.tp14AUC)
p47.AUC

```

```{r}
p45.ldiffsc<-p45.ldiffs
p45.ldiffsc[p45.ldiffsc<0]<-0

p45.tp1AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,2],2))
p45.tp2AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,3],2))
p45.tp3AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,4],2))
p45.tp4AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,5],2))
p45.tp5AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,6],2))
p45.tp6AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,7],2))
p45.tp7AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,8],2))
p45.tp8AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,9],2))
p45.tp9AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,10],2))
p45.tp10AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,11],2))
p45.tp11AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,12],2))
p45.tp12AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,13],2))
p45.tp13AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,14],2))
p45.tp14AUC <- sum(diff(p45.ldiffsc[,1])*rollmean(p45.ldiffsc[,15],2))

p45.AUC<-rbind(p45.tp1AUC,p45.tp2AUC,p45.tp3AUC,p45.tp4AUC,p45.tp5AUC,p45.tp6AUC,p45.tp7AUC,p45.tp8AUC,p45.tp9AUC,p45.tp10AUC,p45.tp11AUC,p45.tp12AUC,p45.tp13AUC,p45.tp14AUC)

p45.AUC
```


```{r}
dates<-c("03-04-19","03-21-19","04-11-19","05-02-19","05-16-19","05-28-19","06-13-19","07-01-19","07-22-19","08-16-19","09-17-19","10-14-19","11-12-19","12-06-19")
auc.df<-data.frame(cbind(dates,p45.AUC[,1],p47.AUC[,1]))

colnames(auc.df)<-c("dates","p45auc","p47auc")
rownames(auc.df)<-c()
Rk.results<-cbind(auc.df,upR45_withrange,lowR45_withrange,R47_withrange)
Rk.results
```

## Making Figure 7

Uses saved data files
ripK_withcomments.csv for quadrat 47, I added columns for incidence
and ripK_longR45.csv for quadrat 45, where I moved from a wide to a long format for quadrat 45 because it was easier to graph and then removed some values (NA) because at some time points there weren't really 2 peak clustering radii

### Start with Quadrat 47
```{r}
ripK<-read.csv("ripK_withcomments.csv")

ripK$dates<-as.Date(ripK$dates,"%m/%d/%y")
ripK
```
```{r}
ripK$scaledI<-ripK$p47.incidence/10
ripK$p47.incidence
ripK$p45.incidence
topdifs<-ripK$top_R47-ripK$R47
bottomdifs<-ripK$R47-ripK$bottom_R47

ripK_p47<-ggplot(ripK)+
  geom_point(mapping=aes(x=dates,y=R47,color=p47.auc))+
  geom_path(mapping=aes(x=dates,y=R47),color="grey")+
  geom_errorbar(aes(x=dates,ymin=R47-bottomdifs, ymax=R47+topdifs,color=p47.auc),width=3)+
  scale_color_viridis(option = "magma",direction=-1,limits = c(0, 17))+
  geom_bar(mapping=aes(x=dates,y=scaledI),stat="identity",fill=" dark grey")+
  scale_y_continuous(name = "Cluster radius (m)", breaks=seq(0,10,1),expand = c(0, 0),limits=c(0,10),sec.axis = sec_axis(~.*10,name ="Disease incidence",breaks=c(0,4,8,12,16,20)))+
  theme(legend.position = "none")+
  theme(panel.background = element_blank())+ 
  theme(text = element_text(family = "Times New Roman"),axis.text.x = element_text(angle = 70, hjust = 1))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x="Survey Dates",color="ABC")+
  scale_x_date( breaks=ripK$dates,date_labels="%d %b %y")+
  guides(size = FALSE)
ripK_p47
```

### Now Quadrat 45
```{r}
longripK_45<-read.csv("ripK_longR45.csv")
longripK_45$dates<-as.Date(longripK_45$dates,"%m/%d/%y")
longripK_45
longripK_45$scaledI<-longripK_45$p45.incidence/10
longripK_45$p45.incidence
ripK_p45<-ggplot(longripK_45)+
  #geom_line(data = longripK_45[!is.na(longripK_45$maxclust), ],aes(x=dates,y=maxclust,group=measure),color="grey")+
  geom_point(mapping=aes(x=dates,y=maxclust,color=p45.auc,group=measure,shape=measure))+
  geom_bar(mapping=aes(x=dates,y=scaledI),stat="identity",fill="dark grey")+
  geom_point(mapping=aes(x=dates,y=maxclust,color=p45.auc,group=measure,shape=measure))+
  geom_path(mapping=aes(x=dates,y=maxclust,group=measure),color="grey")+
  geom_errorbar(aes(x=dates,ymin=maxclust-(maxclust-bottom_R45), ymax=maxclust+(top_R45-maxclust),color=p45.auc),width=3)+
  scale_color_viridis(option = "magma",direction=-1,limits = c(0, 17))+
  scale_y_continuous(name = "Cluster radius (m)", breaks=seq(0,10,1),expand = c(0, 0),limits=c(0,10),sec.axis = sec_axis(~.*10,name ="Disease incidence",breaks=c(0,4,8,12,16,20)))+
  theme(legend.position = "bottom")+
  theme(panel.background = element_blank())+ 
  theme(text = element_text(family = "Times New Roman"),axis.text.x = element_text(angle = 70, hjust = 1))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x="Survey Dates",color="ABC",shape="Quadrat 45 R")+
  scale_x_date( breaks=ripK$dates,date_labels="%d %b %y")+
  guides(size = FALSE)
ripK_p45
```


```{r,fig.height=3.5,fig.width=3}
ripK_p47/ripK_p45 +plot_annotation(tag_levels="a")

```

```{r}
#tiff("Figure7.tiff",width=180, height=220,units="mm",res=300)
ripK_p47/ripK_p45 +plot_annotation(tag_levels="A")
#dev.off()
```

