---
title: "Nearest Neighbors"
output: html_notebook
---
## This notebook includes:

- Data-wrangling for spatial dynamics, using "SCTLD_END_exta.csv"

- Euclidean Distance distribution for all colonies and just diseased colonies (Figure 6)

- Kolmogorov-Smirnoc test for euclidian distances

- Nearest Previously Diseased neighbor monte carlo analysis, as well as supplemental nearest neighbor and nearest diseased neighbor analyses.

```{r,echo=FALSE}
library(zoo) # has rollmeans function for calculating AUC
library(igraph)
library(spdep)
library(spatstat)
library(abind)
library(sciplot)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(ggpubr)
library(vegan)
library(nlme)
library(car)
library(patchwork)
getwd()
```

```{r}
my.data<-read.csv("SCTLD_END_exta.csv") 
##makes a column with factors for if the coral got disease or not throughout the entire survey
#summary(my.data)
my.data$glom<-as.factor(my.data$total_bin) 
my.data$glom<-revalue(my.data$glom, c("0"="Healthy", "1"="SCTLD")) 
```

```{r}
#now keeping data in the wide format
#need just suscpetible species
dis.sps<-my.data%>%
  filter(Sps!="AAGA",Sps!="ACER",Sps!="ATEN",Sps!="EFAS",Sps!="MANG",Sps!="MMEA",Sps!="MYCE",Sps!="OCUL",Sps!="ODIF",Sps!="PAST",Sps!="PDIV",Sps!="PPOR",Sps!="SRAD")%>%
  droplevels()
```

```{r}
#ok, shorten dis.sps dataset to just 10/30/19-12/6/19, becuase didnt see disease until then
## this is the wide format dataset
#colnames(dis.sps)
my.datas<-dis.sps%>%
  dplyr::select(Site,Plot,Sps,coords_x,coords_y,X8.17.18,X10.30.18,X11.9.18,X11.29.18,X12.13.18,X1.4.19,X1.18.19,X2.8.19,X3.4.19,X3.21.19,X4.11.19,X5.2.19,X5.16.19,X5.28.19,X6.13.19,X7.1.19,X7.22.19,X8.16.19,X9.17.19,X10.14.19,X11.12.19,X12.6.19,glom)
#colnames(my.datas)
#refactor data to all have same levels
my.datas$X8.17.18<-factor(my.datas$X8.17.18,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X10.30.18<-factor(my.datas$X10.30.18,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X11.9.18<-factor(my.datas$X11.9.18,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X11.29.18<-factor(my.datas$X11.29.18,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X12.13.18<-factor(my.datas$X12.13.18,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X1.4.19<-factor(my.datas$X1.4.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X1.18.19<-factor(my.datas$X1.18.19,levels=c("Healthy","SCTLD","Dead","Unknown"))

my.datas$X2.8.19<-factor(my.datas$X2.8.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X3.4.19<-factor(my.datas$X3.4.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X3.21.19<-factor(my.datas$X3.21.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X4.11.19<-factor(my.datas$X4.11.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X5.2.19<-factor(my.datas$X5.2.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X5.16.19<-factor(my.datas$X5.16.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X5.28.19<-factor(my.datas$X5.28.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X6.13.19<-factor(my.datas$X6.13.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X7.1.19<-factor(my.datas$X7.1.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X7.22.19<-factor(my.datas$X7.22.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X8.16.19<-factor(my.datas$X8.16.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X9.17.19<-factor(my.datas$X9.17.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X10.14.19<-factor(my.datas$X10.14.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X11.12.19<-factor(my.datas$X11.12.19,levels=c("Healthy","SCTLD","Dead","Unknown"))
my.datas$X12.6.19<-factor(my.datas$X12.6.19,levels=c("Healthy","SCTLD","Dead","Unknown"))

#head(my.datas)
mcavs<-subset(my.datas,subset=Sps=="MCAV")
cnats<-subset(my.datas,subset=Sps=="CNAT")
ofav<-subset(my.datas,subset=Sps=="OFAV")
dsto<-subset(my.datas,subset=Sps=="DSTO")
```

```{r}
my.datas<-dsto
#Split into plots for analysis
s3.p45<-my.datas%>%
  filter(Plot==45)
s3.p47<-my.datas%>%
  filter(Plot==47)

s1.p23<-my.datas%>%
  filter(Plot==23)
s1.p25<-my.datas%>%
  filter(Plot==25)

s2.p27<-my.datas%>%
  filter(Plot==27)
s2.p28<-my.datas%>%
  filter(Plot==28)

```

### Euclidean Distance analysis

#### Get euclidean distances (all colonies and just diseased) for each plot 
```{r}
plotnum<-s1.p23

coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y))
alldists<-as.matrix(dist(coords))
allxy<-t(combn(colnames(alldists), 2))
all.dists<-alldists[allxy]

discoords<-na.omit(subset(plotnum,subset=glom=="SCTLD",select=c(coords_x,coords_y)))
disdists<-as.matrix(dist(discoords))
disxy<-t(combn(colnames(disdists), 2))
dis.dists<-disdists[disxy]

p23.disdists<-dis.dists
p23.alldists<-all.dists
############################
plotnum<-s1.p25

coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y))
alldists<-as.matrix(dist(coords))
allxy<-t(combn(colnames(alldists), 2))
all.dists<-alldists[allxy]

discoords<-na.omit(subset(plotnum,subset=glom=="SCTLD",select=c(coords_x,coords_y)))
disdists<-as.matrix(dist(discoords))
disxy<-t(combn(colnames(disdists), 2))
dis.dists<-disdists[disxy]

p25.disdists<-dis.dists
p25.alldists<-all.dists
##########################
plotnum<-s2.p27

coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y))
alldists<-as.matrix(dist(coords))
allxy<-t(combn(colnames(alldists), 2))
all.dists<-alldists[allxy]

discoords<-na.omit(subset(plotnum,subset=glom=="SCTLD",select=c(coords_x,coords_y)))
disdists<-as.matrix(dist(discoords))
disxy<-t(combn(colnames(disdists), 2))
dis.dists<-disdists[disxy]

p27.disdists<-dis.dists
p27.alldists<-all.dists
##########################
plotnum<-s2.p28

coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y))
alldists<-as.matrix(dist(coords))
allxy<-t(combn(colnames(alldists), 2))
all.dists<-alldists[allxy]

discoords<-na.omit(subset(plotnum,subset=glom=="SCTLD",select=c(coords_x,coords_y)))
disdists<-as.matrix(dist(discoords))
disxy<-t(combn(colnames(disdists), 2))
dis.dists<-disdists[disxy]

p28.disdists<-dis.dists
p28.alldists<-all.dists
##########################
plotnum<-s3.p45

coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y))
alldists<-as.matrix(dist(coords))
allxy<-t(combn(colnames(alldists), 2))
all.dists<-alldists[allxy]

discoords<-na.omit(subset(plotnum,subset=glom=="SCTLD",select=c(coords_x,coords_y)))
disdists<-as.matrix(dist(discoords))
disxy<-t(combn(colnames(disdists), 2))
dis.dists<-disdists[disxy]

p45.disdists<-dis.dists
p45.alldists<-all.dists
##########################
plotnum<-s3.p47

coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y))
alldists<-as.matrix(dist(coords))
allxy<-t(combn(colnames(alldists), 2))
all.dists<-alldists[allxy]

discoords<-na.omit(subset(plotnum,subset=glom=="SCTLD",select=c(coords_x,coords_y)))
disdists<-as.matrix(dist(discoords))
disxy<-t(combn(colnames(disdists), 2))
dis.dists<-disdists[disxy]

p47.disdists<-dis.dists
p47.alldists<-all.dists
```

Combine the distances from the plots
```{r}
all.disdists<-c(p23.disdists,p25.disdists,p27.disdists,p28.disdists,p45.disdists)
all.alldists<-c(p23.alldists,p25.alldists,p27.alldists,p28.alldists,p45.alldists)
all.mcavdisdists<-c(p45.disdists,p47.disdists)
all.mcavalldists<-c(p45.alldists,p47.alldists)
```

#### Figure 6 Euclidean distances
```{r}
all.alldists<-p47.alldists
all.disdists<-p47.disdists
all.med<-median(all.alldists)
dis.med<-median(all.disdists)
all.med
dis.med
bins<-seq(0,14,by=.5)
p<-hist(all.alldists,breaks=bins,plot=FALSE)

#p
binvalsum<-sum(p$counts)
#binvalsum
binvaldivsum<-p$counts/binvalsum
#binvaldivsum
Fvals<-binvaldivsum*100
#Fvals
  
disp<-hist(all.disdists,plot=FALSE,breaks=bins)
#disp
  
dis.binvalsum<-sum(disp$counts)
#dis.binvalsum
dis.binvaldivsum<-disp$counts/dis.binvalsum
#dis.binvaldivsum
dis.Fvals<-dis.binvaldivsum*100
#dis.Fvals
names.5<-c(0,0.5,1,1.5,2,2.5,3,3.5,4,4.5,5,5.5,6,6.5,7,7.5,8,8.5,9,9.5,10,10.5,11,11.5,12,12.5,13,13.5)
names1<-c(0,1,2,3,4,5,6,7,8,9,10,11,12,13)

data.fvals<-rbind(dis.Fvals,Fvals)
#figure
#tiff("Figure6.tiff",width=180, height=150,units="mm",res=300)
par(family="Times New Roman")
bp<-barplot(data.fvals,beside=TRUE,ylim=c(0,10),col=c(grey(.1),grey(.7)),xlab="Distance (m)", ylab="Frequency (%)",las=2,names.arg=names.5,family="Times New Roman")
#title("Euclidean distances between all Surveyed")
legend("topright",cex=.7,bty="n",col=c("black","black"),fill=c("black","grey"),legend=c("Diseased corals","All corals"))
title("D. stokesi")
#abline(v=32,col=grey(.5),lty=2,lwd=2)
#abline(v=29,col="black",lty=2,lwd=2)
axis(1,labels=FALSE,tck=0) 
#text(x=26,y=9.5,"4.66",family="Times New Roman")
#text(x=35,y=9.5,"5.06",family="Times New Roman")
#dev.off()
```

#### Kolmogorov smirnov test
```{r}
my.se<-function(x){
  sd(x,na.rm=TRUE)/sqrt(length(x))
}
ks.test(all.disdists,all.alldists,alternative="g")
#In the ks.test "greater" means that the CDF of 'x' is higher than the CDF of 'y' which means that things like the mean and the median will be smaller values in 'x' than in 'y' if the CDF of 'x' is "greater" than the CDF of 'y'.
mean(all.alldists)
my.se(all.alldists)
mean(all.disdists)
my.se(all.disdists)
```

### Nearest Neightbors analysis

- nearest infected neighbor to diseased coral
- nearest neightbor to diseased coral
- nearest previously infected neighbor to new infected coral

Randomization tests were performed to determine a statistical significance between the realized average nearest infected neighbor distance for all newly infected colonies and that of a null distribution generated using monte carlo methods. The general code procedure is below. It generates a pseudo-p-value that represents the likelihood that the realized average nearest neighbor distance is greater than the null. If the disease follows a contagious model of transmission, we would expect that new infections are closer to previously infected corals, and thus the realized average nearest infected neighbor distance would be smaller than a null distribution (shuffled) average nearest infected neighbor distance. A small pseudo-P-value means that the realized value is smaller than the null distribution, and a pseudo-P-value < 0.5 indicates that the value is significantly smaller and the disease follows a contagious model of transmission.

1.	Determine the distances to the nearest infected neighbor for all newly infected colonies and get the average nearest infected neighbor distance. 
2.	Shuffle the coordinates to randomize which colonies are the infected colonies at each time point
3.	For the shuffled data, determine the distances to the nearest infected neighbor for all newly infected colonies and get the average nearest infected neighbor distance. 
4.	Repeat steps 2 & 3 100 times
5.	Determine the number of times that the realized average nearest infected neighbor distance is greater than the shuffled average nearest infected neighbor distance.
6.	The p value is the value calculated in step 5 divided by the number of simulations

#### Functions needed
```{r}

get_prevIdistmat<-function(df,plot){
  ####newly infected corals
  newly_I<-matrix(0, nrow = nrow(df), ncol = 22) #blank matrix for storing newly infected   corals

  for (i in 1:(22)){ #for each survey time point
    col<-5+i #start with first tp after init tp
  #print(col)
  #print(df[,col])
    prev<-5+i-1
  #print(prev)
    for (j in 1:(nrow(df))){ #for each row in the df
    #print (df[j,col])
      if ( df[j,col]=="SCTLD" ){ #if it's disease
      #print( "found one")
        if (df[j,prev]!="SCTLD"){ # and if it wasnt diseased before
          newly_I[j,i]<-1 #add it to newly infected
        }
      }
    }
  }
#dim(newly_I)


  if(plot==45){
    newly_I[43,15]<-0
    newly_I[40,19]<-0
    newly_I[40,15]<-0
    newly_I[37,15]<-0
  }

  if(plot==47){
    newly_I[2,21]<-0 #for plot 47
  }
  #if (sum(rowSums(newly_I))>nrow(df)){
    #print("ERROR")
    #print(rowSums(newly_I))
  #}

  ##diseased or not matrix
  range<-ncol(df)
  df_dis_short<-df[,6:range] #shorten the dataframe to just the month data
#nrow(df_dis_short)
  dis_or_not<-matrix(0,nrow=nrow(df_dis_short),ncol=ncol(df_dis_short)) #set up blank matrix to store the dis or not as 1s and 0s
  for (i in 1:nrow(dis_or_not)){ #over each row
    for (j in 1:ncol(dis_or_not)){ #over each column
      if (df_dis_short[i,j]=="SCTLD"){ #if not healthy
        dis_or_not[i,j]<-1 #label as diseased / 1 in new matrix
      }
    }
  }
#dim(dis_or_not)

###### get distances to previously infected corals mat where timepoints are columns and corals are rows
  distmat<-as.matrix(dist(cbind(df$coords_x,df$coords_y)))
  dist_to_nearest_prev_I<-c() #set up blank list to track nearest infected neighbor to newly infected
  track_dists_newlyI <- matrix(NA,nrow(df)) #blank matrix for storing all distances between old inf and new inf
#dim(newly_I)
  for (i in 1:nrow(newly_I)){ #for each row in newly_I, which is each coral
  #print(i)
    for (j in 1:ncol(newly_I)){ #for each column in newly_I, which is each time point
    #print(j)
      coral<-newly_I[i,j] #choose a coral at the timepoint
      if (coral>0){ #if the coral is newly infected (1)
      #print("new inf coral")
      #print(i)
      #print(j)
        mat_prev_inf<-as.matrix(dis_or_not[,j]) #choose the column before coral's column to get 1 inf and 0 not infected; because dis_or_not has all time point columns j is the timepoint before the timepoint being pulled from the newly_I
      #mat_prev_inf is a matrix of 1 column (the timepoint before the coral timepoint we are considering) where 0 is not infected and 1 is infected
      #print(mat_prev_inf)
        dists_between_newI_coral_and_all<-distmat[,i] #choose the ith column of the distance matrix which will be the column of the coral we are looking at as newly infected --> this will be a numeric class
        all_dists_between_newI_coral_and_prevI_coral<-dists_between_newI_coral_and_all*mat_prev_inf #row will be 0 if the prev time point it was not infected. it will be the distance if it was previously infected
        track_dists_newlyI<-cbind(track_dists_newlyI,all_dists_between_newI_coral_and_prevI_coral) #track all of these distances
      #dists_between_newI_coral_and_prevI_coral<-all_dists_between_newI_coral_and_prevI_coral[which(rowSums(all_dists_between_newI_coral_and_prevI_coral) > 0),]  #get rid of zeros 
      #print(dists_between_newI_coral_and_prevI_coral)
      #dist_to_nearest_prev_I<-append(dist_to_nearest_prev_I,min(dists_between_newI_coral_and_prevI_coral)) #append the minimum dist to prev inf to the list
      }
    }
  }
return(track_dists_newlyI)
}
#get_NPDNmat(df,plot)

nn_prevI_to_newI<-function(dists){
  #dists is a matrix where each column is a coral and it's distances to previously infected corals (each row is a dif coral and the val is dist, 0 if not infected before the colyumn coral)
  nn_newIprevI<-c()
  for (i in 2:ncol(dists)){
    coralcol<-dists[,i] #choose one colony to look at
    #print(coralcol)
    coralcol<-as.matrix(coralcol) #turn it into a matrix
    coralcol<-coralcol[which(rowSums(coralcol)>0)] #get rid of zeros
    nn_newIprevI<-append(nn_newIprevI,min(coralcol)) #get the nearest neighbor distance
  }
  return(nn_newIprevI)
}

```

### nearest neighbors for all corals
```{r}
s3.p45n<-na.omit(s3.p45)
s3.p45n$NN<-nndist(s3.p45n$coords_x,s3.p45n$coords_y)

s3.p47n<-na.omit(s3.p47)[-141,]
s3.p47n$NN<-nndist(s3.p47n$coords_x,s3.p47n$coords_y)

s1.p23n<-na.omit(s1.p23)
s1.p23n$NN<-nndist(s1.p23n$coords_x,s1.p23n$coords_y)

s1.p25n<-na.omit(s1.p25)
s1.p25n$NN<-nndist(s1.p25n$coords_x,s1.p25n$coords_y)

s2.p27n<-na.omit(s2.p27)
s2.p27n$NN<-nndist(s2.p27n$coords_x,s2.p27n$coords_y)

s2.p28n<-na.omit(s2.p28)
s2.p28n$NN<-nndist(s2.p28n$coords_x,s2.p28n$coords_y)


```
nearest diseased neighbors for diseased corals
```{r}

s3.p45d<-na.omit(subset(s3.p45n,subset=glom=="SCTLD"))
s3.p45d$NDN<-nndist(s3.p45d$coords_x,s3.p45d$coords_y)

s3.p47d<-na.omit(subset(s3.p47n,subset=glom=="SCTLD"))
s3.p47d$NDN<-nndist(s3.p47d$coords_x,s3.p47d$coords_y)

s1.p23d<-na.omit(subset(s1.p23n,subset=glom=="SCTLD"))
s1.p23d$NDN<-nndist(s1.p23d$coords_x,s1.p23d$coords_y)

s1.p25d<-na.omit(subset(s1.p25n,subset=glom=="SCTLD"))
s1.p25d$NDN<-nndist(s1.p25d$coords_x,s1.p25d$coords_y)

s2.p27d<-na.omit(subset(s2.p27n,subset=glom=="SCTLD"))
s2.p27d$NDN<-nndist(s2.p27d$coords_x,s2.p27d$coords_y)

s2.p28d<-na.omit(subset(s2.p28n,subset=glom=="SCTLD"))
s2.p28d$NDN<-nndist(s2.p28d$coords_x,s2.p28d$coords_y)


```

#### Do multiple simulations for monte carlo randomization stat method.

```{r}
plotdisdata<-list(s3.p47d,s3.p45d,s2.p27d,s2.p28d,s1.p23d,s1.p25d)
plotdisdata<-list(s3.p47d,s3.p45d)
plotnumbers<-list(47,45,27,28,23,25)
plotnumbers<-list(47,45)

sites<-list(s3.p47,s3.p45)
bins<-seq(0,14,by=.5)

all.nnprevI<-c()
all.nn<-c()
all.ndn<-c()
all.2ndn<-c()
all.3ndn<-c()
all.all.dists<-c()
all.dis.dists<-c()
for (i in 1:6){
  #nearest previously infected coral to newly infected coral
  plot<-plotnumbers[i]
  plotdis<-as.data.frame(plotdisdata[i])
  df<-plotdis#[,-c(28,29,30,31)]
  dists_prevInewI<-get_prevIdistmat(df,plot)
  nnprevI<-nn_prevI_to_newI(dists_prevInewI)
  nnprevI<-nnprevI[!is.infinite(nnprevI)] #remove infs
  all.nnprevI<-c(all.nnprevI,nnprevI)
  
  #nearest neighbors and euclidean distances for all colonies 
  plotnum<-as.data.frame(sites[i])
  coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y))
  all.nn<-c(all.nn,nndist(coords)) #nearest neighbors
  alldists<-as.matrix(dist(coords))
  allxy<-t(combn(colnames(alldists), 2))
  all.dists<-alldists[allxy]
  all.all.dists<-c(all.all.dists,all.dists) #euclidean distances
  
  #nearest diseased neighbors for diseased colonies and euclidean distances btwn dis corals
  discoords<-na.omit(subset(plotnum,subset=glom=="SCTLD",select=c(coords_x,coords_y)))
  disdists<-as.matrix(dist(discoords))
  all.ndn<-c(all.ndn,nndist(discoords)) #nearest diseased neighbors
  all.2ndn<-c(all.2ndn,nndist(discoords,k=2)) #second nearest
  all.3ndn<-c(all.3ndn,nndist(discoords,k=3)) #third nearest

  disxy<-t(combn(colnames(disdists), 2))
  dis.dists<-disdists[disxy]
  all.dis.dists<-c(all.dis.dists,dis.dists) #euclidean distances
}
real.anpin<-mean(all.nnprevI)
real.ann<-mean(all.nn)
real.andn<-mean(all.ndn)
real.a2ndn<-mean(all.2ndn)
real.a3ndn<-mean(all.3ndn)
real.adist<-mean(all.all.dists)
real.adisdist<-mean(all.dis.dists)


#random simulations
sims<-1000
rand.anpin<-c()
rand.andn<-c()
rand.a2ndn<-c()
rand.a3ndn<-c()
rand.adisdist<-c()
for (i in 1:sims){
  rand.nnprevI<-c()
  rand.ndn<-c()
  rand.2ndn<-c()
  rand.3ndn<-c()
  rand.dis.dists<-c()
  for (j in 1:6){
    plot<-plotnumbers[j]
    plotnum<-as.data.frame(sites[j]) #all colonies
    plotdis<-as.data.frame(plotdisdata[j]) #diseased colonies
    numdis<-nrow(plotdis) #number of diseased colonies
    
    coords<-na.omit(cbind(plotnum$coords_x,plotnum$coords_y)) #just get coords of all cols
    randcoords<-coords[sample(nrow(coords), numdis), ] #randomly sample n diseased corals
    #replace diseased corals coords with random coords
    plotdis$coords_x<-randcoords[,1] 
    plotdis$coords_y<-randcoords[,2]
    
    #redo the nearest neighbor analysis
    df<-plotdis[,-c(28,29,30,31)]
    dists_prevInewI<-get_prevIdistmat(df,plot)
    nnprevI<-nn_prevI_to_newI(dists_prevInewI)
    nnprevI<-nnprevI[!is.infinite(nnprevI)]
    rand.nnprevI<-c(rand.nnprevI,nnprevI)
    
    #now for randomized nn and distances, not just previously infected
    rand.ndn<-c(rand.ndn,nndist(randcoords)) #nearest diseased neighbors
    rand.2ndn<-c(rand.2ndn,nndist(randcoords,k=2)) #second nearest
    rand.3ndn<-c(rand.3ndn,nndist(randcoords,k=3)) #third nearest
    randdists<-as.matrix(dist(randcoords))
    randxy<-t(combn(colnames(randdists),2))
    rand.dists<-randdists[randxy]
    rand.dis.dists<-c(rand.dis.dists,rand.dists) #euclidean distances
  }
  rand.anpin<-c(rand.anpin,mean(rand.nnprevI))
  rand.andn<-c(rand.andn,mean(rand.ndn))
  rand.a2ndn<-c(rand.a2ndn,mean(rand.2ndn))
  rand.a3ndn<-c(rand.a3ndn,mean(rand.3ndn))
  rand.adisdist<-c(rand.adisdist,mean(rand.dis.dists))
}

N.extreme_anpin <- sum(rand.anpin < real.anpin)
p_anpin <- N.extreme_anpin / (sims)
print("p_anpin")
p_anpin

N.extreme_andn <- sum(rand.andn < real.andn)
p_andn <- N.extreme_andn/ (sims)
print("p_andn")
p_andn

N.extreme_a2ndn <- sum(rand.a2ndn < real.a2ndn)
p_a2ndn <- N.extreme_a2ndn/ (sims)
print("p_a2ndn")
p_a2ndn

N.extreme_a3ndn <- sum(rand.a3ndn < real.a3ndn)
p_a3ndn <- N.extreme_a3ndn/ (sims)
print("p_a3ndn")
p_a3ndn


N.extreme_add <- sum(rand.adisdist < real.adisdist)
p_add <- N.extreme_add/ (sims)
print("p_add")
p_add

```

```{r}
#all euclidean dist
print("all colonies euclidean distances mean and SE")
mean(all.all.dists)
my.se(all.all.dists)
#dis dists
print("Just Disease colonies euclidean distances mean and SE")
mean(all.dis.dists)
my.se(all.dis.dists)
#real.anpin<-mean(all.nnprevI)
#real.ann<-mean(all.nn)
#real.andn<-mean(all.ndn)
#real.a2ndn<-mean(all.2ndn)
#real.a3ndn<-mean(all.3ndn)
#all nn
print("all colonies nearest neighbors distances mean and SE")
mean(all.nn)
my.se(all.nn)
# ndn
print("nearest diseased neighbors distances mean and SE")
mean(all.ndn)
my.se(all.ndn)
# rand ndn
print("random nearest diseased neighbors distances mean and SE")
mean(rand.ndn)
my.se(rand.ndn)
# npin
print("nearest previously diseased neighbor distances mean and SE")
mean(all.nnprevI)
my.se(all.nnprevI)
# rand npdn
print("randomized nearest previously diseased neighbor distances mean and SE")
mean(rand.nnprevI)
my.se(rand.nnprevI)
```



